---
title: "Examples from Chapter 2"
author: "Miles D. Williams"
date: '2022-06-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get Packages

```{r}
library(tidyverse)
library(socviz)
```

## The Work We'll Be Doing

Most of what we'll be doing with Healy's book will entail working with a series of datasets. Here's a very small one from the `socviz` liberary:

```{r}
titanic
class(titanic)
```

The `titanic` dataset contains two numeric and two non-numeric columns. So that our dataset provides more explicit information about the variables, we can convert the data frame into a `tibble`:

```{r}
titanic_tb <- as_tibble(titanic)
titanic_tb
```

To see inside objects we can use the `str()` function:

```{r}
str(titanic_tb)
```

On occasion, we'll need to read in data, either stored locally in a format such as .csv or a remote file like below:

```{r}
url <- "https://cdn.rawgit.com/kjhealy/viz-organdata/master/organdonation.csv"
organs <- read_csv(url)
```

The message produced when using `read_csv()` (which we use to read in .csv files --- *shocker!*) reports information about the variables and how they've been read in in the console. There are other data types we can read in as well using variations on the `read_[type]` syntax.

## A First Visual

Healy has us use the Gapminder dataset, which is already available in R, to perform our first vizualization:

```{r}
library(gapminder)
gapminder
```

This dataset provides info about several countries over time. Specifically, about life expectancy, population, and GDP per capita.

We can make a scatter plot with some of these variables as follows:

```{r}
p <- ggplot(data = gapminder) +
  aes(
    x = gdpPercap,
    y = lifeExp
  )
p + geom_point()
```
